CSD, Disjoint components and trayectory detection

### 2019-07-19 ArticuloCSDyCM/Paper03.tex
tabla de trayectorias

The necessity to create such analysis came originally from the study of
electro-physiological data obtained from the rat's hippocampus, using
a very dense MEA, the BioCAM of 3Brain \cite{BioCAM}. 
The highly layered CA structure of the hippocampus provides strong, discernible patterns of Sources and Sinks. Following the pattern of activity using Center of Mass analysis directly from the data, though, proves unfruitful, due to the non-convex nature of the somatic layer. Using CSD and disconnected component tracking over the data provided the right conceptual framework to follow the displacement of active areas during highly active events, such as epileptic-like activity that was pharmacological
or traumatically induced. The proposed analysis would not produce vector averages over different simultaneously active areas. On the contrary, it would separate them and track them independently. So the most violent epileptic burst provided us with the data necessary to prove the robustness of the method.
All the experimental data comes from a series of experiments
performed on slices of the rats hippocampus.

@@

The figure \ref{lfpcsdcm}  compares the representation of the activity in
the LFP, CSD and CM spaces. The CSD
representation provides better spatial localization of active sites, but the
CM can serve as better dynamical tracker of action. The figure illustrates
that point by choosing the time interval of response in CA3 after the stimulus.
CM representation permits us to track the activity with less
interference from the saturated region. The small, clearly localized centers of
mass separate clearly from the big patch over the saturated region.
In the figure \ref{lfpcsdcmsub} we repeat the procedure but focusing in
the r.o.i., which includes CA3 str. pyr. We take advantage of the
detailed frame to show the borders of the disconnected sets that
we have chosen for the CM analysis in the right column.

@@

\begin{figure}
 \includegraphics[width=0.9\textwidth]{Figuras/CMVariousTimes01.pdf}
  \caption{ The CM analysis over the same images as figure \ref{lfpcsdcmsub} C. Some trajectories
    are calculated before the 4ms time stamp. The triangular mark shows the later point of the
    trajectory, either the end or the last calculated position until that moment. The circular
    mark is the start of the trajectory. We can see that activity moves form CA3 distal to
    proximal in alternating waves of action (hiperpolarization, firing and repolarization).
  }\label{cmevocada}
\end{figure}

\begin{figure}
  \includegraphics[width=0.9\textwidth]{./Figuras/trayevocadasub01.pdf}
  \caption{The evoked trajectories in the r.o.i. We begin to track them
    at the 0.9 ms mark, after letting the saturation of the stimulus recede.
    At 3ms a very clear sink trajectory is moving retrogradly towards CA2,
    sheepherd by two shorter source trayectories. At 5.6 a depolarization
    trajectory starts, following almost the same path and presenting also its
    two companions, ending around the 7.7ms timestamp. No significant
  trajectories are observed after the 10ms mark. }\label{trayevo}
\end{figure}

>>>>

The necessity to create such analysis came originally from the study of
electro-physiological data obtained from the rat's hippocampus, using
a very dense MEA, the BioCAM of 3Brain \cite{BioCAM}. 
The highly layered CA structure of the hippocampus provides strong, discernible patterns of Sources and Sinks. Following the paths of activity using Center of Mass analysis directly from the data, though, proves unfruitful, due to the non-convex nature of the somatic layer. Using CSD and disconnected component tracking over the data provided the right conceptual framework to follow the displacement of active areas during highly active events, such as epileptic-like activity that was pharmacological
or traumatically induced. The proposed analysis would not produce vector averages over different simultaneously active areas. On the contrary, it would separate them and track them independently. So the most violent epileptic burst provided us with the data necessary to prove the robustness of the method.
All the experimental data comes from a series of experiments
performed on slices of the rats hippocampus.

@@

The figure \ref{lfpcsdcm}  compares the representation of the activity in
the LFP, CSD and CM spaces. The CSD
representation provides better spatial localization of active sites, and the
CM serves as better dynamical tracker. The figure illustrates
that point by choosing the time interval of response in CA3 after the stimulus.
CM representation permits us to track the activity with less
interference from the saturated region. The small, clearly localized centers of
mass separate clearly from the big patch over the saturated region.
In the figure \ref{lfpcsdcmsub} we repeat the procedure but focusing in
the r.o.i., which includes CA3 str. pyr. We take advantage of the
detailed frame to show the borders of the disconnected sets that
we have chosen for the CM analysis in the right column.

@@

\begin{figure}[h]
 \includegraphics[width=0.9\textwidth]{Figuras/CMVariousTimes01.pdf}
  \caption{ The CM analysis over the same images as figure \ref{lfpcsdcmsub} C. Some trajectories
    are calculated before the 4ms time stamp. The triangular mark shows the later point of the
    trajectory, either the end or the last calculated position until that moment. The circular
    mark is the start of the trajectory. We can see that activity moves form CA3 distal to
    proximal in alternating waves of action (hiperpolarization, firing and repolarization).
  }\label{cmevocada}
\end{figure}

\begin{figure}[h]
  \includegraphics[width=0.9\textwidth]{./Figuras/trayevocadasub01.pdf}
  \caption{The evoked trajectories in the r.o.i. We begin to track them
    at the 0.9 ms mark, after letting the saturation of the stimulus recede.
    At 3ms a very clear sink trajectory is moving retrogradly towards CA2,
    sheepherd by two shorter source trayectories. At 5.6 a depolarization
    trajectory starts, following almost the same path and presenting also its
    two companions, ending around the 7.7ms timestamp. No significant
  trajectories are observed after the 10ms mark. }\label{trayevo}
\end{figure}
